---
title: "p8105_hw3_mbc2178"
author: "Melvin Coleman"
date: "2022-10-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's load in the libraries needed to perform data manipulation and cleaning for this assignment.
```{r message = FALSE}
library(tidyverse)
library(dplyr)
library(readr)
```

## Problem 1 

Let's load the `instacart` data from the `p8105.datasets` using the following:
```{r}
library(p8105.datasets)
data("instacart") 
```

The `instacart` dataset contains `r nrow(instacart)` observation 
and `r ncol(instacart)` variables. The variables that exist in this dataset include 
`r ls(instacart)`. 


## Problem 2 

Let's import the `accel_data` dataset to R. This dataset contains five weeks of 
accelorometer data collected on a 63 year old male with BMI 25, who was admitted 
to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed 
with congestive hear failure (CHF).

We will perform some tidying and wrangle the data before proceeding to perform 
any analysis. First, we will consider using `janitor::clean_names()` to clean 
the names of the variables in the dataset. Next, we will use the `pivot_longer`
function to create a new variable that contains the activity count names and 
another variable that contains the activity counts for each munute of a 24-hr 
day starting at midnight.

```{r}
accel_df =
  read_csv(file = "data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
```



